Linear analysis of the vertical shear instability in astrophysical disks

### 2015-01-21 numerical.tex
regular update

\subsection{Effect of thermal relaxation}
We now introduce thermal relaxation. Here we consider a slightly
different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as used by
\cite{nelson13}. Fig. \ref{relax_growth_num} plots the growth rates
of the fundamental mode function of $k_x$. %corrugation mode
% We first obtain the eigenvalues for the
% fundamental mode for $k_xH_\mathrm{iso}=0$, then vary parameters
% slowly to obtain eigenvalues for other values of $k_x$ or $\beta$.

As above, Fig. \ref{relax_growth_num} compares well with
Fig. \ref{relax_disp_fig} except at $k_xH_\mathrm{iso}\lesssim 1$
where the low-frequency approximation fails. The non-smoothness in the
curve for $\beta=0.063$ at $k_xH_\mathrm{iso}\gtrsim50$ was found to be
associated with perturbations near the vertical boundaries. This is
likely attributable to the free surface boundary condition 
(cf. regularity condition at infinity in \S\ref{disp_relax}).

@@

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag_bloop2}
  \caption{Growth rates of the fundamental VSI mode in
    nearly-vertically isothermal disks ($\Gamma=1.011$) with
    $\epsilon=0.05$ and $q=-1$, with thermal relaxation timescale
    $\beta$. The eigenfrequencies are calculated by numerically
    solving the full linear eigenvalue problem. This plot is to
    be compared with Fig. \ref{relax_disp_fig}. \label{relax_growth_num}}   
\end{figure}

We visiualize the effect of thermal relaxation in
Fig. \ref{relax_eigenW_num} by comparing the eigenfunction $W$ for
$\beta=0.01$ and $\beta=0.125$, the latter is close to the critical
value beyond which the isothermal fundamental VSI is stabilized
(\S\ref{iso_vsi_beta_crit}).

@@

\citeauthor{nelson13} find a thermal relaxation timescale of
$\beta\gtrsim 0.6$ stabilized the VSI for this disk model in non-linear
hydrodynamic simulations. This is consistent with our numerical
results.

>>>>

\subsection{Effect of thermal relaxation}
We now introduce thermal relaxation. Here we consider a slightly
different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as
adopted in \cite{nelson13}. Fig. \ref{relax_growth_num} plots the
growth rates of the fundamental mode function of $k_x$. %corrugation
                                %mode 
% We first obtain the eigenvalues for the
% fundamental mode for $k_xH_\mathrm{iso}=0$, then vary parameters
% slowly to obtain eigenvalues for other values of $k_x$ or $\beta$.

As above, Fig. \ref{relax_growth_num} is largely consistent with 
Fig. \ref{relax_disp_fig} except for $k_xH_\mathrm{iso}\lesssim 1$
where the low-frequency approximation fails. The non-smoothness in the
curve for $\beta=0.063$ at $k_xH_\mathrm{iso}\sim50$ was found to be
associated with large perturbations near the vertical boundaries. This is
likely attributable to the free surface boundary condition imposed in
the numerical calculation (cf. regularity condition at infinity in
\S\ref{disp_relax}).

@@

\begin{figure}
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_imag_bloop}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag_bloop2}
  \caption{Growth rates of the fundamental VSI mode in
    nearly-vertically isothermal disks ($\Gamma=1.011$) with
    $\epsilon=0.05$ and $q=-1$, with thermal relaxation timescale
    $\beta$. The eigenfrequencies are calculated by numerically
    solving the full linear eigenvalue problem. \label{relax_growth_num}}   
\end{figure}

We visualize the effect of thermal relaxation in 
Fig. \ref{relax_eigenW_num} by comparing the eigenfunction $W$ for
$\beta=0.01$ and $\beta=0.125$. The latter value is close to the critical
value beyond which the fundamental isothermal VSI is stabilized
(\S\ref{iso_vsi_beta_crit}).  Increasing $\beta$ restricts 
the region in which $W\sim z$ closer to the midplane, with oscillatory
behavior emerging at the boundaries, where buoyancy first becomes
important.

@@

%\citeauthor{nelson13} find a thermal relaxation timescale of
% $\beta\gtrsim 0.6$ stabilized the VSI for this disk model in non-linear
% hydrodynamic simulations. This is consistent with our numerical
% results.

\subsection{Vertically non-isothermal disks}
%gamma = Gamma = 1.4, vary bcool

%gamma = 2.0, Gamma = 1.4, vary bcool
